\documentclass[11pt]{article}

\usepackage{times}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage[psamsfonts]{amssymb}
\usepackage{latexsym}
\usepackage{color}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{amstext}
\usepackage{url}
\usepackage{epsfig}
\usepackage{bm}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
}
\usepackage{mathtools}
 
\def\Kset{\mathbb{K}}
\def\Nset{\mathbb{N}}
\def\Qset{\mathbb{Q}}
\def\Rset{\mathbb{R}}
\def\Sset{\mathbb{S}}
\def\Zset{\mathbb{Z}}
\def\squareforqed{\hbox{\rlap{$\sqcap$}$\sqcup$}}
\def\qed{\ifmmode\squareforqed\else{\unskip\nobreak\hfil
\penalty50\hskip1em\null\nobreak\hfil\squareforqed
\parfillskip=0pt\finalhyphendemerits=0\endgraf}\fi}

\DeclareMathOperator*{\E}{\rm E}
\DeclareMathOperator*{\argmax}{\rm argmax}
\DeclareMathOperator*{\argmin}{\rm argmin}
\DeclareMathOperator{\sgn}{sign}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\last}{last}
\DeclareMathOperator{\sign}{\sgn}
\DeclareMathOperator{\diag}{diag}
\providecommand{\abs}[1]{\lvert#1\rvert}
\providecommand{\norm}[1]{\lVert#1\rVert}
\def\vcdim{\textnormal{VCdim}}

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\newcommand{\cX}{{\mathcal X}}
\newcommand{\cY}{{\mathcal Y}}
\newcommand{\cA}{{\mathcal A}}
\newcommand{\ignore}[1]{}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\bd}{\begin{description}}
\newcommand{\ed}{\end{description}}
\newcommand{\h}{\widehat}
\newcommand{\e}{\epsilon}
\newcommand{\mat}[1]{{\mathbf #1}}
\newcommand{\R}{\mat{R}}
\newcommand{\0}{\mat{0}}
\newcommand{\M}{\mat{M}}
\newcommand{\D}{\mat{D}}
\renewcommand{\r}{\mat{r}}
\newcommand{\x}{\mat{x}}
\renewcommand{\u}{\mat{u}}
\renewcommand{\v}{\mat{v}}
\newcommand{\w}{\mat{w}}
\renewcommand{\H}{\text{0}}
\newcommand{\T}{\text{1}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\xxi}{{\boldsymbol \xi}}
\newcommand{\ssigma}{{\boldsymbol \sigma}}
\newcommand{\Alpha}{{\boldsymbol \alpha}}
\newcommand{\tts}{\tt \small}
\newcommand{\hint}{\emph{hint}}
\newcommand{\matr}[1]{\bm{#1}}     % ISO complying version
\newcommand{\vect}[1]{\bm{#1}}     % ISO complying version
\renewcommand{\labelitemi}{$\bullet$}
\renewcommand{\labelitemii}{$\cdot$}
\renewcommand{\labelitemiii}{$\diamond$}
\renewcommand{\labelitemiv}{$\ast$}

\newenvironment{solution}{\vspace{.25cm}\noindent{\it Solution:}}{}

\begin{document}

\noindent DS-GA 1008:\\
Deep Learning, Spring 2019\\
Homework Assignment 2 \\
Yves Greatti - yg390\\
\href{https://www.overleaf.com/project/5c7316d634ff9f6576d02da0}{Read-Only Link to tex file}


\section*{1. Fundamentals}
\subsection*{1.1. Convolution}
Table \ref{tab1} depicts two matrices. The one on the left represents a $5\times 5$ single-channel image $\matr{A}$. The one on the right
 represents a $3\times 3$ convolution kernel $\matr{B}$. 
\begin{itemize}
    \item[(a)] What is the dimensionality of the output if we forward propagate the image over the given convolution kernel with no padding and stride of 1? \\
    
     \smallskip
    If you apply the convolution kernel with no padding and a stride of 1 over the input image  $\matr{A}$ we obtain an output image of size (3,3).
    This is illustrated in the animations  from \href{https://github.com/vdumoulin/conv_arithmetic} {"A technical report on convolution arithmetic in the context of deep learning"}
    but in our case, we have no padding and a stride of one.
    
    \item[(b)] Give a general formula of the output width $O$ in terms of the input width $I$, kernel width $K$, stride $S$, and padding $P$ (both in the beginning and in the end). Note that the same formula holds for the height. Make sure that your answer in part (a) is consistent with your formula. \\
    
    \smallskip
    The padding $P$ being at the beginning and at the end, and having a stride of $S$, we obtain a general formula for the output width which will be similar to the one for the height :
    \[
    	O  = \floor*{\frac{I + 2 * P - K}{S} + 1}
    \]
    For   $5\times 5$ image $\matr{A}$, and $3\times 3$  convolution kernel $\matr{B}$, with P=0 and S=1, we have $O_{\text{width}} =  \floor*{\frac{5 + 2 * 0 - 3}{1}  + 1} = 3 = O_{\text{height}}$
    
    \item[(c)] Compute the output $\matr{C}$ of forward propagating the image over the
given convolution kernel. Assume that the bias term of the convolution is zero.\\

	\smallskip
	If we apply the kernel $\matr{B}$ to the upper corner of size   $3\times 3$ of the image $\matr{A}$, we have: 
	$4 \times 3 + 5 \times 3 + 2 \times 3 + 3 \times 5 +  3 \times 5 + 2 \times 5 + 4 \times 2 + 3 \times 4 + 4 \times 3 = 109$
	Similarly we can find the entries for the rest of the output $\matr{C}$ :

	    \begin{table}[!ht]
		    \centering
    		$\matr{C}$ = \begin{tabular}{|c|c|c|} 
    			\hline
			    109 & 92 & 72 \\ \hline 
		        108 & 85 & 74 \\ \hline
       			110 & 74 & 79 \\ \hline 
   		 \end{tabular}
    	  \end{table}

    \item[(d)] Suppose the gradient backpropagated from the layers above this layer is a $3\times 3$ matrix of all 1s. Write the value of the gradient (w.r.t.~the input image) backpropagated out of this layer.\\
    
    \smallskip
    Using the chain rule, we have:
   \begin{align*}
    	\frac{\partial E}{ \partial  \matr{A}_{ij}} &= \sum_{k,l=1,3} \frac{\partial E}{\partial \matr{C}_{kl}}   \frac{\partial \matr{C}_{kl}} {\partial  \matr{A}_{ij}} \\
	&= \sum_{k,l=1,3}  \frac{\partial \matr{C}_{kl}} {\partial  \matr{A}_{ij}}
    \end{align*}
    Expanding this, we obtain:
   \begin{gather*}
   	\frac{\partial E}{ \partial  \matr{A}_{11}} = \frac{\partial \matr{C}_{11}} {\partial  \matr{A}_{11}} =   \matr{B}_{11} \\
   	\frac{\partial E}{ \partial  \matr{A}_{12}} = \frac{\partial \matr{C}_{11}} {\partial  \matr{A}_{12}} + \frac{\partial \matr{C}_{12}} {\partial  \matr{A}_{12}} =    \matr{B}_{12} + \matr{B}_{11} \\
	\frac{\partial E}{ \partial  \matr{A}_{13}} = \frac{\partial \matr{C}_{11}} {\partial  \matr{A}_{13}} + \frac{\partial \matr{C}_{12}} {\partial  \matr{A}_{13}} + \frac{\partial \matr{C}_{13}} {\partial  \matr{A}_{13}}  =   \matr{B}_{13} + \matr{B}_{12} + \matr{B}_{11} \\
	\vdots
    \end{gather*}
    If we keep expanding each term we realize that each term is the result of a "full" convolution between the loss gradient $\frac{\partial E}{\partial \matr{C}}$, which is a matrix of all 1s, and a filter $\matr{B}$ (\href{https://www.mathworks.com/help/matlab/ref/conv2.html}{MATLAB conv2}).
    We have the result:
    
    	    \begin{table}[!ht]
		    \centering
    		 $\frac{\partial E}{\partial \matr{A}}$ = \begin{tabular}{|c|c|c|c|c|}  
    			\hline
    			 4  & 7  & 10 &  6 & 3 \\ \hline 
		         9  & 17 & 25 & 16 & 8 \\ \hline
		        11  & 23 & 34 & 23 & 11 \\ \hline 
		         7  & 16 & 24 & 17 & 8 \\ \hline
		         2  & 6  &  9 &  7 & 3 \\ \hline
    		\end{tabular}\hspace{1cm}
        	 \end{table}

\end{itemize}

 
\begin{table}[!ht]
    \centering
    $\matr{A}$ =  \begin{tabular}{|c|c|c|c|c|} 
    \hline
       4 & 5 & 2 & 2 & 1 \\ \hline 
       3 & 3 & 2 & 2 & 4 \\ \hline
       4 & 3 & 4 & 1 & 1 \\ \hline 
       5 & 1 & 4 & 1 & 2 \\ \hline
       5 & 1 & 3 & 1 & 4 \\ \hline
    \end{tabular}\hspace{1cm}
    $\matr{B}$ = \begin{tabular}{|c|c|c|} 
    \hline
       4 & 3 & 3 \\ \hline 
       5 & 5 & 5 \\ \hline
       2 & 4 & 3 \\ \hline 
    \end{tabular}
    \caption{Image Matrix ($5\times 5$) and a convolution kernel ($3\times 3$).}\label{tab1}   
\end{table}

 Hint: You are given that $\frac{\partial E}{\partial \matr{C}_{ij}} = 1$ for some scalar error $E$ and $i,j\in\{1,2,3\}$. You need to compute $\frac{\partial E}{\partial \matr{A}_{ij}}$ for  $i,j\in\{1, \ldots, 5\}$. The chain rule should help!

\subsection*{1.2. Pooling}
The pooling is a technique for sub-sampling and comes in different flavors, for example max-pooling, average pooling, LP-pooling. 
\begin{itemize}
    \item[(a)] List the \texttt{torch.nn} modules for the 2D versions of these pooling techniques and read on what they do.\\
    \medskip
    \\
    The 2D pooling layers are
	\begin{itemize}
		\item MaxPool2d\\
		In regular max pooling, you downsize an input set by taking the maximum value of smaller N x N subsections of the set (often 2x2), and try to reduce the set by a factor of N, where N is an integer.
		\item AvgPool2d\\
		Performs the average pooling on the input.
		\item FractionalMaxPool2d\\
		Fractional max pooling is slightly different than regular max pooling.  Fractional max pooling, as you might expect from the word "fractional", means that the overall reduction ratio N does not have to be an integer. The sizes of the pooling regions are generated randomly but are fairly uniform.
		\item LPPool2d\\
		LP-pooling performs a 2D power-average pooling over an input signal. 
		 Each LP-pooling defines a spherical shape in a non-Euclidean space whose metrics is defined  by  the lp-norm.
		\item AdaptiveMaxPool2d\\
		Adaptive max pooling is a max operation per channel.
		The tensor before the average pooling is supposed to have as many channels as your model has classification categories.
		\item AdaptiveAvgPool2d\\
		Adaptive average pooling is similar to adaptive max-pooling where the operation is an average instead of the maximum.
	\end{itemize}
    
    
    \item[(b)] Denote the $k$-th input feature maps to a pooling module as $\matr{X}^k \in \mathbb{R}^{H_{\textrm{in}}\times W_{\textrm{in}}} $ where $H_{\textrm{in}}$ and $W_{\text{in}}$ represent the input height and width, respectively. Let $\matr{Y}^k \in \mathbb{R}^{H_{\text{out}}\times W_{\textrm{out}}}$ denote the $k$-th output feature map of the module where $H_{\textrm{out}}$ and $W_{\textrm{out}}$ represent the output height and width, respectively. Let $S^{k}_{i,j}$ be a list of the indexes of elements in the sub-region of $X^k $ used for generating $\matr{Y}^k_{i,j}$, the $(i,j)$-th entry of $\matr{Y}^{k}$. 
    Using this notation, give formulas for $\matr{Y}^k_{i,j} $ from three pooling modules.\\
    
	\begin{itemize}
		\item MaxPool2d
		\[
			\matr{Y}^k_{i,j} = \max \{ X^k_{l,m}| (l,m) \in  S^{k}_{i,j} \}
		\]
		\item AvgPool2d
		\[
			\matr{Y}^k_{i,j} = \frac{1}{|  S^{k}_{i,j} |} \sum_{(l,m) \in  S^{k}_{i,j}} X^k_{l,m} ~~  \text{where } |  S^{k}_{i,j} | \text{: number of elements (l,m) in } S^{k}_{i,j}
		\]	
		\item LPPool2d
		\[
			\matr{Y}^k_{i,j} =   (\sum_{(l,m) \in  S^{k}_{i,j}} (X^k_{l,m})^p)^{\frac{1}{p}}
		\]	
		Note that PyTorch defines LPPool2d as $(\sum_{x \in X} x^p)^{\frac{1}{p}}$
		
	\end{itemize}
        
    
    \item[(c)] Write out the result of applying a max-pooling module with kernel size of 2 and stride of 1 to $\matr{C}$ from Part 1.1.\\
    \smallskip
    Applying max-pooling with a kernel of size 2 and stride 1 to , we obtain the following output:
    
    	    \begin{table}[!ht]
		    \centering
    		$\text{max-pooling} (\matr{C})$ = \begin{tabular}{|c|c|} 
    			\hline
			109 & 92 \\ \hline 
		        110 & 85 \\ \hline
   		 \end{tabular}
    	  \end{table}

    
    \item[(d)] Show how and why max-pooling and average pooling can be expressed in terms of LP-pooling.\\
    \smallskip
    	\begin{itemize}
		\item Max-pooling:
		    Short answer: if we apply max-pooling to the feature map $X$, 
    		    since $ \|X\|_{\infty}  = \max_{x_i \text{ component of X}}  \{ |x_i| \} = \lim_{p \to \infty} \|X\|_p$ , thus LP-pooling is max-pooling for $p \to \infty$.
		\item Average-pooling:
		Reusing the notations in question 1.2.b, then for p=1:
		\begin{align*}
			\text{LP-pooling} (S^k_{i,j})  &=  \sum_{(l,m) \in  S^{k}_{i,j}} X^k_{l,m} \\
						&= |  S^{k}_{i,j} | \text{ Average-pooling}(S^k_{i,j})
		\end{align*}
		   
 	\end{itemize}   
\end{itemize}

\end{document}
